Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2021-07-20 paper/paper.tex
Modified border effect to be effective from r, not 2r - neater and perfectly sensible

We implement a border effect based on gradient following by assuming that
there is some other molecular signal which acts on all branches near the
boundary of the tectal tissue. For a branch with position $(x,y)$, $\mathbf{B}_b$ is
given by:
%
\begin{equation}
B_{b,x} = \begin{cases}
        2r-x    & x<2r \\
        1-2r-x    & x>1-2r
\end{cases}
B_{b,y} = \begin{cases}
        2r-y    & y<2r \\
        1-2r-y    & y>1-2r
\end{cases}
\end{equation}
%
Thus $\mathbf{B}_b$ is equivalent to the action of a repulsive signalling
molecule expressed around the border of the tectum, whose expression increases
quadratically outside the tectum.

@@

\begin{figure}
\includegraphics[width=\linewidth]{./images/j4_ee_wt.png}
\caption{Gradient following model with axon-axon interaction}
\label{f:wt}
\end{figure}

We first present the result of running the model above, which assumes that
axonal growth cones express graded levels of receptors which interact with
ligands expressed as gradients on the tectum and on the surface of other
growth cones to guide their movement. We make the following additional
assumptions: i) All receptor-ligand interactions are repulsive, as for
EphA-ephrinA coupling \citep{nakamoto_topographically_1996}; and ii) tectal
ligand expressions vary exponentially along a given axis, following
Eq.\,\ref{e:tecexp}.

@@

>>>>

We implement a border effect based on gradient following by assuming that
there is some other molecular signal which acts on all branches near the
boundary of the tectal tissue. For a branch with position $(x,y)$, $\mathbf{B}_b$ is
given by:
%
\begin{equation}
B_{b,x} = \begin{cases}
        r-x      & x<r \\
        1-r-x    & x>1-r
\end{cases}
B_{b,y} = \begin{cases}
        r-y      & y<r \\
        1-r-y    & y>1-r
\end{cases}
\end{equation}
%
Thus $\mathbf{B}_b$ is equivalent to the action of a repulsive signalling
molecule expressed around the border of the tectum, whose expression increases
quadratically outside the tectum and affects any branch touching (or outside)
the boundary

@@

\begin{figure}
\includegraphics[width=\linewidth]{./images/j4_ee_wt.png}
\caption{Gradient following model with axon-axon interactions.}
\label{f:wt}
\end{figure}

To enable a simulation it is necessary to make the assumptions, in addition to
those implicit in the description of the model above, that i) All
receptor-ligand interactions are repulsive, as for EphA-ephrinA
coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}; and ii)
all tectal ligand expressions vary exponentially along a given axis, following
Eq.\,\ref{e:tecexp}. The result of running this system are presented in
Fig.\,\ref{f:wt}.
%
Fig.\,\ref{f:wt}A shows a dual-colour map of retinal ganglion cell positions
on the square retina, with red encoding position on the $x$ axis (temporal to
nasal) and green encoding position on the $y$ axis (ventral to
dorsal). Fig.\,\ref{f:wt}B presents a fishnet plot of the expected location on
the tectal surface of each RGC at the end of the development process. Thus,
the `red' RGC from the ventro-nasal corner should map to the caudo-medial
corner of the tectum and the `green' RGC from the dorso-temporal corner should
find its way to the rostro-lateral corner on the tectum. The lines between the
dots of Fig.\,\ref{f:wt}B indicate the adjacency relationships between
cells. Thus, colour in each panel indicates a branch's location of origin on
the retina.
%
Fig.\,\ref{f:wt}C shows the initial state of the system, with branches
stochastically arranged around the rostral edge of the tectum. Each dot
represents the centroid of $n=8$ branches per RGC axon. Figs.\,\ref{f:wt}D-E
show the development of the system with time, with Fig.\,\ref{f:wt}E showing

@@

The model described above assumes that branches express graded levels of
receptors which interact with ligands expressed as gradients on the tectum,
the tectal border and on the surface of other branches, and that only these
interactions guide their movement.
